\documentclass[8pt]{article}

\usepackage{fullpage}
\usepackage[margin=.7in]{geometry}
\usepackage{epic}
\usepackage{eepic}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{ragged2e}
\usepackage[parfill]{parskip}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\proof}[1]{
{\noindent {\it Proof.} {#1} \rule{2mm}{2mm} \vskip \belowdisplayskip}
}

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}


\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{claim}[lemma]{Claim}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}

\begin{document}
\hfill \small{\today} \\
\setlength{\fboxrule}{.5mm}\setlength{\fboxsep}{1.2mm}
\newlength{\boxlength}\setlength{\boxlength}{\textwidth}
\addtolength{\boxlength}{-4mm}
\begin{center}\framebox{\parbox{\boxlength}{\bf
\center{CS 577 - Homework 5}
\center{Sejal Chauhan, Vinothkumar Siddharth, Mihir Shete}
}}\end{center}
\vspace{5mm}

\section{Graded written problem}

\textbf{Input:} In a city there are $n$ bus drivers. There are also $n$ morning bus routes and $n$ afternoon bus routes, each with various lengths. Each driver is assigned one morning route
and one evening route. For any driver, if his total route length for a day exceeds $d$, he has to
be paid overtime for every hour after the first $d$ hours at a fixed rate per hour.
\\ \\
\textbf{Output:} Assign one morning route and one evening route to each bus driver so that the total overtime amount that the city authority has to pay is minimized.

\subsection{Algorithm}
Our greedy algorithm begins by sorting the $n$ morning bus routes in reverse order (Longest route first) and the $n$ afternoon bus routes in order. To find a set $G\{(m,a): \forall m \in M \land \forall a \in A \}$, where M is set of all morning routes and A is the set of all evening routes, we will use the following \textit{strategy}:
\begin{enumerate}
    \item Starting from the first morning(largest) route we get the smallest available afternoon route and pair them.
\end{enumerate}


\begin{algorithm}
\caption{Pseudocode of our greedy strategy}\label{euclid}
\begin{algorithmic}[1]
\Procedure{Find-All-Pairs}{$M$, $A$}

\State $i \leftarrow 0$

\While{$i \textless \left\vert{M}\right\vert$}
        \State Pair $M[i]$, $A[i]$
        \State $i \leftarrow i + 1$
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Exchange Argument}
Let $r$ be all the route-pairs in a solution that are greater than $d$ and let $l_i$ be the length of the route-pair . $p(G) = \sum_{r}(l_i - d)$ for our greedy solution $G$, similarly for an optimal Solution $S$ $p(S) = \sum_{r}(l_i - d)$

If, $G = S$ then clearly $p(G) = p(S)$.

Otherwise, $G \neq S$. So, there must be some 2 routes in $G$ and $S$, such that the morning routes are the same but the paired afternoon routes are different. More formally, if ($M_G(i)$, $A_G(i)$), ($M_G(j)$, $A_G(j)$) are 2 pairs in G and ($M_S(k)$, $A_S(k)$), ($M_S(l)$, $A_S(l)$) are 2 pairs in S, then
$$M_G(i) = M_S(k) \And M_G(j) = M_S(l)$$
but,
$$A_G(i) \neq A_S(k) \And A_G(j) \neq A_S(l)$$ \\
Let, $M_{ik} = M_G(i) = M_S(k)$ and $M_{jl} = M_G(j) = M_S(l)$
Now, consider the following scenarios in $S$:
\begin{description}
    \item[$M_{ik} \textgreater M_{jl}$] \hfill \\
        Let $S'$ denote a solution which would have selected the ($M_{ik}, A_S(l)$) and ($M_{jl}, A_S(k)$) pair contrary to $S$. $S'$ is similar to $G$ in a way that $G$ would also have chosen the same pairing as $S'$. So, given the constraint on morning routes $S'$ will choose the afternoon route first for $M_{ik}$ and it will pair with $A_S(l)$ if:

    \begin{enumerate}
        \item \textbf{$A_S(l) \textless A_S(k)$ and $M_{ik} + A_S(l) \leq d$} \hfill \\
            For the above constraints $M_{jl} + A_S(k)$ can be less than, more than or equal to $d$, so lets analyze all the conditions:
            Here, if $$M_{jl} + A_S(k) \textgreater d$$ it follows that $$M_{ik} + A_S(k) \textgreater d$$ But, $$p(S') = C + M_{jl} + A_S(k) - d$$
            $$p(S) = C + M_{ik} + A_S(k) - d$$ $\because M_{jl} \textless M_{ik}$ it implies that $p(S') \textless p(S)$ \\ \\
            $M_{jl} + A_S(k) \leq d$ then the overtime introduced by the swap is $0$ and so $p(S') \leq p(S) $ \\ \\


        \item \textbf{$A_S(l) \textless A_S(k)$ and $M_{ik} + A_S(l) \textgreater d$} \hfill \\
            For the above constraints $M_{jl} + A_S(k)$ can be less than, more than or equal to $d$, so lets analyze all the conditions:
            Here, if $M_{jl} + A_S(k) \textgreater d$ and $M_{jl} + A_S(l) \textgreater d$ it is trivial to see that $p(S') = p(S)$ since the extra hours introduced for both the selections will be $M_{ik} + M_{jl} + A_S(k) + A_S(l) - 2$. \\ \\
            But if $M_{jl} + A_S(k) \textgreater d$ and $M_{jl} + A_S(l) \leq d$
            $$p(S) = M_{ik} + A_S(k) - d$$
            $$p(S') = M_{ik} + A_S(l) + M_{jl} + A_S(k) - 2d = p(S) + M_{jl} + A_S(l) - d \leq p(S) \because M_{jl} + A_S(l) \leq d$$\\

            If, $M_{jl} + A_S(k) \leq d$ then it implies that $M_{jl} + A_S(l) \leq d$ from the constraints. $$p(S') =  C + M_{ik} + A_S(l) - d$$ And $$p(S) = C + M_{ik} + A_S(k) - d$$
            And since $A_S(l) \textless A_S(k)$ it is clear that $p(S') \textless p(S)$
    \end{enumerate}
\end{description}

From the above analysis it is clear that $p(S') \leq p(S)$ whenever the order is changed according to greedy strategy. Hence, our greedy strategy will provide an optimal solution.

\subsection{Running Time Analysis}
To sort get the routes in correct order for our greedy algorithm to process we will have to spend $O(n\log{n})$ time. The \textbf{Find-All-Pairs} procedure will iterate over all $n$ elements in $M$ and all the operations in these iterations are constant time. So overall runtime is:


$$O(n \log{n} + c*n) = O(n \log{n})$$

\end{document}
