\documentclass[8pt]{article}

\usepackage{fullpage}
\usepackage[margin=.7in]{geometry}
\usepackage{epic}
\usepackage{eepic}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{ragged2e}
\usepackage[parfill]{parskip}

\newcommand{\proof}[1]{
{\noindent {\it Proof.} {#1} \rule{2mm}{2mm} \vskip \belowdisplayskip}
}

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}


\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{claim}[lemma]{Claim}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}

\begin{document}
\hfill \small{\today} \\
\setlength{\fboxrule}{.5mm}\setlength{\fboxsep}{1.2mm}
\newlength{\boxlength}\setlength{\boxlength}{\textwidth}
\addtolength{\boxlength}{-4mm}
\begin{center}\framebox{\parbox{\boxlength}{\bf
\center{CS 577 - Homework 5}
\center{Sejal Chauhan, Vinothkumar Siddharth, Mihir Shete}
}}\end{center}
\vspace{5mm}

\section{Graded written problem}

\textbf{Input:} In a city there are $n$ bus drivers. There are also $n$ morning bus routes and $n$ afternoon bus routes, each with various lengths. Each driver is assigned one morning route
and one evening route. For any driver, if his total route length for a day exceeds $d$, he has to
be paid overtime for every hour after the first $d$ hours at a fixed rate per hour.
\\ \\
\textbf{Output:} Assign one morning route and one evening route to each bus driver so that the total overtime amount that the city authority has to pay is minimized.

\subsection{Algorithm}
Our greedy algorithm begins by sorting the $n$ morning bus routes in reverse order (Longest route first) and the $n$ afternoon bus routes in order. To find a set $G\{(m,a): \forall m \in M \land \forall a \in A \}$, where M is set of all morning routes and A is the set of all evening routes, we will use the following \textit{strategy}:
\begin{enumerate}
    \item Starting from the first morning route we try to find the \textit{largest} afternoon route such that their sum is less than or equal to $d$. 
    \item If there is no such afternoon route such that the sum of the morning and afternoon route is less than or equal to $d$ then we will pair the smallest available afternoon route with the given morning route.
    \item Rule \textit{2.} Also implies that if morning route is greater than $d$ we will pair it with the smallest available afternoon route.
\end{enumerate}

Consider the following morning and evening routes ordered as per our algorithm: \\ \\
$Morning:$ \{$7,4,3,1$\} $Afternoon:$ \{$2,3,6,9$\} $d:8$ \\ \\
In the above example we will start from the largest morning route $7$, we can see that the smallest available afternoon route and the largest morning route add to a value greater than $d$, since there is no available route in the afternoon with which we can pair the largest route in the morning to get the total duration less than $d$ we will pair $7$ and $2$ (i.e the smallest available afternoon route). Similarly, we will pair ($4$, $3$), ($3$, $6$) and ($1$, $9$) according to our greedy algorithm.

If we do time complexity analysis of the \textit{strategy} of pairing the routes we can see that in the worst case we will have to compare the $1^{st}$ route in the morning to $n$ routes in the afternoon, the $2^{nd}$ route in the morning to $n - 1$ routes in the afternoon and so on. Hence, our time complexity will be $O(n^2)$. To better the running time of our strategy we will use the following algorithm (a variant of binary search). This algorithm will take $key$ as one of the inputs where $key = d - m_i$, $m_i$ is the route duratio of the $i^{th}$ morning trip and will return the best match (as decribed in our strategy) from the array $A$ of size $\left\vert{A}\right\vert$ which is the array of ordered Afternoon trips.

\begin{algorithm}
\caption{Find-Pair}\label{euclid}
\begin{algorithmic}[1]
\Procedure{Find-Pair}{$key$, $A$, $left$, $right$}

\If{$left \textgreater right$}
    \Return left
\EndIf

\State $mid \leftarrow \floor{(right + left) / 2}$
    \If{$key \textless A[mid]$}
        \State \Return \Call{Find-Pair}{$key$, $A$, $left$, $mid - 1$}
    \ElsIf{$key \textgreater A[mid]$}
        \State \Return \Call{Find-Pair}{$key$, $A$, $mid + 1$, $right$}
    \ElsIf{$key = A[k]$}
        \State \Return k
    \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}

If $key = d - m_i$ is present in the array $A$ then \textbf{Find-Pair} will return an index $k$ such that $A[k] = key$ and we can directly pair this trip of length $A[k]$ with trip $m_i$. if $key$ is not present and $A[k]$ \textless $key$ then we will also pair $k$ with $m_i$, but if $A[k] \textgreater m_i$ then we will pair $m_1$ with $A[k-1]$ if $k \textgreater 0$. The runtime of \textbf{Find-Pair} is $log(n)$ because we are divind the array in half for every iteration and using this algorithm we can complete our strategy on ordered arrays in $O(nlog(n))$ time.

Let $M$ be the array of all ordered morning trips and $A$ be the array reresenting all ordered afternoon trips, our \textit{strategy} will be represented by the following Pseudocode:

\begin{algorithm}
\caption{Pseudocode of our greedy stratgey}\label{euclid}
\begin{algorithmic}[1]
\Procedure{Find-All-Pairs}{$M$, $A$}

\State $i \leftarrow 0$

\While{$i \textless \left\vert{A}\right\vert$}
    \If {$M[i] \geq d$}
        \State Pair $M[i]$, $A[0]$
        \State Delete $A[0]$ from $A$
        \State continue
    \EndIf
    \State $key \leftarrow d - M[i]$
    \State $index \leftarrow \Call{Find-Pair}{key, A, 0, \left\vert{A}\right\vert}$
    \If {$A[index] \leq key$}
        \State Pair $M[i]$, $A[index]$
        \State Delete $A[index]$ from $A$
    \Else
        \If{$index = 0$}
            \State Pair $M[i]$, $A[index]$
            \State Delete $A[index]$ from $A$
        \Else
            \State Pair $M[i]$, $A[index - 1]$
            \State Delete $A[index - 1]$ from $A$
        \EndIf
    \EndIf
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Exchange Argument}
Let $r$ be all the route-pairs in a solution that are greater than $d$ and let $l_i$ be the length of the route-pair . $p(G) = \sum_{r}(l_i - d)$ for our greedy solution $G$, similarly for an optimal Solution $S$ $p(S) = \sum_{r}(l_i - d)$

If, $G = S$ then clearly $p(G) = p(S)$.

Otherwise, $G \neq S$. So, there must be some 2 routes in $G$ and $S$, such that the morning routes are the same but the paired afternoon routes are different. More formally, if ($M_G(i)$, $A_G(i)$), ($M_G(j)$, $A_G(j)$) are 2 pairs in G and ($M_S(k)$, $A_S(k)$), ($M_S(l)$, $A_S(l)$) are 2 pairs in S, then
$$M_G(i) = M_S(k) \And M_G(j) = M_S(l)$$
but,
$$A_G(i) \neq A_S(k) \And A_G(j) \neq A_S(l)$$ \\
Let, $M_{ik} = M_G(i) = M_S(k)$ and $M_{jl} = M_G(j) = M_S(l)$
Now, consider the following scenarios in $S$:
\begin{description}
    \item[$M_{ik} \textgreater M_{jl}$] \hfill \\
        Let $S'$ denote a solution which would have selected the ($M_{ik}, A_S(l)$) and ($M_{jl}, A_S(k)$) pair contrary to $S$. $S'$ is similar to $G$ in a way that $G$ would also have chosen the same pairing as $S'$. So, given the constraint on morning routes $S'$ will choose the afternoon route first for $M_{ik}$ and it will pair with $A_S(l)$ if:

    \begin{enumerate}
        \item \textbf{$A_S(l) \textgreater A_S(k)$ and $M_{ik} + A_S(l) \leq d$} \hfill \\
            In this scenario, it is trivial to see that $M_{jl} + A_S(k) \leq d$ as both morning and afternoon trips are lesser than their counterparts i.e $M_{jl} \textless M_{ik}$ and $A_S(k) \textless A_S(l)$. So, the choice of pair for $S'$ does not yield any overtime. Hence, $p(S') \leq p(S)$ \\

        \item \textbf{$A_S(l) \textless A_S(k)$ and $M_{ik} + A_S(l) \textgreater d$} \hfill \\
            For the above constraints $M_{jl} + A_S(k)$ can be less than, more than or equal to d, so lets aanalyze all the conditions:
            Here, if $M_{jl} + A_S(k) \textgreater d$ it is trivial to see that $p(S') = p(S)$ since the extra hours for both the selections will be $2d - M_{ik} - M_{jl} - A_S(k) - A_S(l)$. \\ \\
            If, $M_{jl} + A_S(k) \leq d$ then it implies that $M_{jl} + A_S(l) \leq d$ from the constraints. $$p(S') =  M_{ik} + A_S(l) - d$$ And $$p(S) = M_{ik} + A_S(k) - d$$
            And since $A_S(l) \textless A_S(k)$ it is clear that $p(S') \textless p(S)$
    \end{enumerate}
\end{description}

From the above analysis it is clear that $p(S') \leq p(S)$ whenever the order is changed according to greedy strategy. Hence, our greedy strategy will provide an optimal solution.

\subsection{Running Time Analysis}
To sort get the routes in correct order for our greedy algorithm to process we will have to spend $O(n\log{n})$ time. The \textbf{Find-All-Pairs} procedure will iterate over all $n$ elements in $M$ and all the operations in these itereations are constant time except \textbf{Find-Pair}.


\end{document}
